---
title: "Nobel laureates"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
# packages
library(learnr)
library(gradethis)
library(tidyverse)

# use gradethis for checking
tutorial_options(exercise.timelimit = 60, exercise.checker = gradethis::grade_learnr)

# make nobel data frame globally available
nobel <- read_csv("https://bit.ly/nobel-laureates")

# hide non-exercise code chunks
knitr::opts_chunk$set(echo = FALSE)
```

---

**This is a demo tutorial created for the "Feed me back: Tales of automated feedback in R and Python" breakout session at [eCOTS 2020](https://www.causeweb.org/cause/ecots/ecots20). See [here](https://bit.ly/feed-me-back) for slides and full set of materials from the breakout session.**

---

## Introduction

In January 2017, Buzzfeed published an article on why Nobel laureates show 
immigration is so important for American science. You can read the article 
[here](https://www.buzzfeednews.com/article/peteraldhous/immigration-and-science). 
In the article they show that while most living Nobel laureates in the 
sciences are based in the US, many of them were born in other countries. 
This is one reason why scientific leaders say that immigration is vital for 
progress. We will work with the data from this article to recreate some of 
their visualizations as well as explore new questions.

### Learning goals

- Manipulate and transform data to prepare it for visualization.
- Recreate visualizations.
- Summarise data.
- Get more practice working as a team.

### Packages

We'll use the **tidyverse** package for this analysis. Run the following code 
to load this package.

```{r load-packages, exercise=TRUE}
library(tidyverse)
```

## Data

The dataset for this assignment can be found as a csv file at https://bit.ly/nobel-laureates. 
You can read it in directly using the `read_csv()` function.

```{r load-data, exercise=TRUE}
nobel <- read_csv("https://bit.ly/nobel-laureates")
```

The variable descriptions are as follows:

- `id`: ID number
- `firstname`: First name of laureate
- `surname`: Surname 
- `year`: Year prize won
- `category`: Category of prize
- `affiliation`: Affiliation of laureate
- `city`: City of laureate in prize year
- `country`: Country of laureate in prize year
- `born_date`: Birth date of laureate
- `died_date`: Death date of laureate
- `gender`: Gender of laureate
- `born_city`: City where laureate was born
- `born_country`: Country where laureate was born
- `born_country_code`: Code of country where laureate was born
- `died_city`: City where laureate died
- `died_country`: Country where laureate died
- `died_country_code`: Code of country where laureate died
- `overall_motivation`: Overall motivation for recognition
- `share`: Number of other winners award is shared with
- `motivation`: Motivation for recognition

In a few cases the name of the city/country changed after laureate was given (e.g. in 1975 Bosnia and Herzegovina was called Austria-Hungary). In these cases the variables below reflect a different name than their counterparts without the suffix `_original`.

- `born_country_original`: Original country where laureate was born
- `born_city_original`: Original city where laureate was born
- `died_country_original`: Original country where laureate died
- `died_city_original`: Original city where laureate died
- `city_original`: Original city where laureate lived at the time of winning the award
- `country_original`: Original country where laureate lived at the time of winning the award

## Explore

### Glimpse

Let's start with a peek at your data.

```{r glimpse-data, exercise=TRUE}
glimpse(nobel)
```

```{r quiz}
quiz(
  question("How many observations and how many variables are in the dataset?",
    answer("26 observations, 935 variables", 
           message = "Hint: Each row is an observation and each column is a variable."),
    answer("935 observations, 26 variables", 
           correct = TRUE,
           message = "The dataset includes information on 935 nobel winners, and 26 attributes for each one."),
    allow_retry = TRUE
  )
)
```

### Filter the data

For the rest of the analysis we'll focus on nobel winners who fit the following criteria:

- laureates for whom `country` is available
- laureates who are people as opposed to organizations (organizations are 
denoted with `"org"` as their `gender`)
- laureates who are still alive (their `died_date` is `NA`)
  
Filter the `nobel` dataset for the conditions specified above. Call the new dataset `nobel_living`.

```{r nobel-filter, exercise=TRUE}
___ <- nobel %>%
  ___(
    ___(country),
    gender ___ "org",
    ___(died_date)
  )
```

```{r nobel-filter-solution}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date)
  )
```

```{r nobel-filter-check}
grade_code("You've successfully filtered the data.")
```

### Check your results

How many nobel laureates fit the criteria? Use the code chunk below to obtain 
this number.

```{r nobel-living-nrow-prepare}
nobel_living <- nobel %>%
  filter(
    !is.na(country),
    gender != "org",
    is.na(died_date)
  )
```

```{r nobel-living-nrow, exercise=TRUE, exercise.setup="nobel-living-nrow-prepare"}

```

```{r nobel-living-nrow-check}
grade_result(
  pass_if(~ identical(.result, 228L), "There are 228 nobel laureates that fit the criteria."),
  fail_if(~ identical(.result, 26L), "Each row represents a nobel laureate. Did you calculate the number of columns instead of rows?"),
  fail_if(~ TRUE, "Not quite. Each row represents a nobel laureate. Try calculating the number of rows.")
)
```
